### 2020-06-23 manuscript/internship_report.Rmd
change abstract

abstract: |
  Eye-tracking allows researchers to infer cognitive processes from eye movements that are classified into distinct events. Parsing the events is typically done by algorithms that transform, filter, classify, and merge raw data samples. Previous algorithms have successfully used hidden Markov models (HMMs) for classification but still inhere weaknesses. I developed gazeHMM, an algorithm that improves on previous HMM algorithms. The development was guided by the question of whether HMMs are useful at describing eye movements and whether they improve the event classification. A simulation study showed that gazeHMM successfully recovered HMM parameters and hidden state sequences with only few exceptions. When applied to benchmark data, model comparisons for gazeHMM with different events did not match theoretical expectations. Compared to other algorithms, gazeHMM improved the event classification which was assessed by the similarity to human coding, especially for dynamic stimuli. The classification of smooth pursuits was found to be inadequate and is thus only recommended for exploration. Otherwise, I consider gazeHMM as appropriate for practical applications. In sum, gazeHMM improves classification but is not accurately describing eye movements, yet. I identify several explanations and discuss potential remedies for this result.

>>>>

abstract: |
  Eye-tracking allows researchers to infer cognitive processes from eye movements that are classified into distinct events. Parsing the events is typically done by algorithms that transform, filter, classify, and merge raw data samples. Previous algorithms have successfully used hidden Markov models (HMMs) for classification but still inhere weaknesses. Therefore, I developed gazeHMM, an HMM algorithm that has no critical parameters to be set by users, does not require human coded data as input, and classifies fixations, saccades, PSOs, and smooth pursuits. The development was guided by the question of whether HMMs are useful at describing eye movements and whether they improve the event classification. I evaluated gazeHMM in a simulation study and on benchmark data. The simulation study showed that gazeHMM successfully recovered HMM parameters and hidden state sequences. Critical exceptions for good recovery were adding a smooth pursuit like state to gazeHMM and noisy data. For benchmark data, model comparisons with gazeHMM yielded preferred models with more states than expected. I assessed the classification performance of gazeHMM compared to other algorithms by the agreement to human event coding. Here, gazeHMM improved the event classification but did not outperform all other algorithms in most cases. Both the simulation study and benchmark application showed poor classification performance for smooth pursuits. Thus, I advice to classify smooth pursuits only for exploration and recommend gazeHMM with three events for application. Future HMM algorithms could use covariates to better capture eye movement processes and explicitly model event durations to improve the classification of smooth pursuits.

### 2020-06-23 manuscript/report_appendix.Rmd
change abstract

(ref:tab-example-trans) Transition probabability estimates for example data from Andersson et al. (2017)
(ref:tab-example-trans-note) Rows indicate from which state and columns into which state the model transitions.

@@

tab.exp.tr <- cbind(1:4, matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[5:20], 2), 4, 4, byrow = T))

apa_table(tab.exp.tr,
          col.names = c("From\nstate", as.character(1:4)),
          col_spanners = list("To state" = c(2,5)),
          caption = "(ref:tab-example-trans)",
          note = "(ref:tab-example-trans-note)")

@@

(ref:tab-example-resp) Response parameter estimates for example data from Andersson et al. (2017)
(ref:tab-example-resp-note) Con = concentration parameter of the von-Mises distribution.

@@

tab.exp.resp <- cbind(1:4, matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[21:44], 2), 4, 6, byrow = T))

tab.exp.resp[1, c(5,6)] <- NA

apa_table(tab.exp.resp,
          col.names = c("State", "Shape (vel)", "Scale (vel)", "Shape (acc)", "Scale (acc)", "Mean", "Con"),
          format.args = list(na_string = "-"),
          caption = "(ref:tab-example-resp)",
          note = "(ref:tab-example-resp-note)")

>>>>

(ref:tab-example-trans) Transition probabability estimates for example data from Andersson et al. (2017)

@@

tab.exp.tr <- cbind(as.character(1:4), matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[5:20], 3), 4, 4, byrow = T))

apa_table(tab.exp.tr,
          col.names = c("From", as.character(1:4)),
          col_spanners = list("To" = c(2,5)),
          caption = "(ref:tab-example-trans)")

@@

(ref:tab-example-resp) Response parameter estimates for example data from Andersson et al. (2017)

@@

tab.exp.resp <- cbind(as.character(1:4), matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[21:44], 2), 4, 6, byrow = T))

tab.exp.resp[1, c(6,7)] <- NA

apa_table(tab.exp.resp,
          col.names = c("State", "Shape (vel)", "Scale (vel)", "Shape (acc)", "Scale (acc)", "Mean", "Con"),
          format.args = list(na_string = "-"),
          caption = "(ref:tab-example-resp)")
